---
title: "High dimensional time series analysis"
date: "robjhyndman.com/hdtsa"
author: "5. Forecast reconciliation"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: ../header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 60)
library(fpp3)
library(broom)
# Create nice R figures with minimal margins
# in landscape format suitable for slides and papers
savepdf <- function(file, width=16, height=10)
{
  fname <<- paste("figs/",file,".pdf",sep="")
  pdf(fname, width=width/2.54, height=height/2.54, pointsize=10, bg='white')
  par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
}
# Crop pdf to remove all white space
endpdf <- function()
{
  #dev.off()
  crop::dev.off.crop(fname)
}
```

# Forecast reconciliation

## Australian tourism
\full{regions1_with_labels}

## Australian tourism

\begin{textblock}{10}(1,1.5)\small
\begin{block}{}
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Quarterly data on visitor night from 1998:Q1 -- 2013:Q4
    \item From: \textit{National Visitor Survey}, based on annual interviews of 120,000 Australians aged 15+, collected by Tourism Research Australia.
    \item Split by 7 states, 27 zones and 76 regions (a geographical hierarchy)
    \item Also split by purpose of travel
      \begin{itemize}
        \item Holiday
        \item Visiting friends and relatives (VFR)
        \item Business
        \item Other
      \end{itemize}
    \item 304 bottom-level series
  \end{itemize}
\end{block}
\end{textblock}

## Spectacle sales

\placefig{1}{1.4}{width=9.5cm}{spectacles.jpg}
\vspace*{4.8cm}

 * Monthly UK sales data from 2000 -- 2014
 * Provided by a large spectacle manufacturer
 * Split by brand (26), gender (3), price range (6), materials (4), and stores (600)
 * About 1 million bottom-level series

## Hierarchical time series
\fontsize{13}{14}\sf

A \alert{\textbf{hierarchical time series}} is a collection of several time series that are linked together in a hierarchical structure.

\begin{minipage}{9.6cm}
\begin{block}{}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=33mm,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AA}}
   child {node {AB}}
   child {node {AC}}
 }
 child {node {B}
   child {node {BA}}
   child {node {BB}}
   child {node {BC}}
 }
 child {node {C}
   child {node {CA}}
   child {node {CB}}
   child {node {CC}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}\vspace*{-0.2cm}

 * Tourism demand by state and region

## Grouped time series
\fontsize{13}{14}\sf

A \alert{\textbf{grouped time series}} is a collection of time series that can be grouped together in a number of non-hierarchical ways.

\begin{minipage}{9.2cm}
\begin{block}{}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
 };
\end{tikzpicture}\hspace*{1cm}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {X}
   child {node {AX}}
   child {node {BX}}
 }
 child {node {Y}
   child {node {AY}}
   child {node {BY}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}

 * Spectacle sales by brand, gender, stores, etc.
 * Tourism by state and purpose of travel

## The problem
\fontsize{13}{14}\sf

\begin{alertblock}{}
\begin{enumerate}\tightlist
 \item How to forecast time series at all nodes such that the forecasts add up in the same way as the original data?
 \item Can we exploit relationships between the series to improve the forecasts?
\end{enumerate}
\end{alertblock}\pause

### The solution

1. Forecast all series at all levels of aggregation using an automatic forecasting algorithm.\newline (e.g., `ETS`, `ARIMA`, ...)
2. Reconcile the resulting forecasts so they add up correctly using least squares optimization (i.e., find closest reconciled forecasts to the original forecasts).
3. This is available using `reconcile()`.

## Hierarchical and grouped time series

Every collection of time series with aggregation constraints can be written as
\begin{block}{}
\centerline{$\by_{t}=\bS\bm{b}_{t}$}
\end{block}
where

 * $\by_t$ is a vector of all series at time $t$
 * $\bm{b}_t$ is a vector of the most disaggregated series at time $t$
 * $\bS$ is a ``summing matrix'' containing the aggregation constraints.




## Hierarchical time series

\begin{minipage}{4cm}\vspace*{0.2cm}
\begin{block}{}\centering
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.3cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
 }
 child {node {B}
 }
 child {node {C}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\only<2->{\begin{textblock}{6.3}(6,1)\small
\begin{itemize}\itemsep=0cm\parskip=0cm
\item[\color{white} $ y_{t}: $] observed aggregate of all series at time
$t$.
\item[\color{white} $ y_{X,t}: $] observation on series $X$ at time $t$.
\item[\color{white} $ \bm{b}_{t}: $] vector of all series at bottom level
in time $t$.
\end{itemize}
\end{textblock}}\vspace*{0.6cm}
\only<3->{
$\bY_{t}= \begin{pmatrix}
  y_{t}\\
  y_{A,t}\\
  y_{B,t}\\
  y_{C,t}
  \end{pmatrix} = \only<3>{\hspace*{0.01cm}\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}\only<4->{{\color{orange}\underbrace{\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}_{\bS}}}\only<3>{\hspace*{0.08cm}}\only<3>{\hspace*{-0.1cm}\begin{pmatrix}Y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}\rule{0cm}{1.6cm}
                \only<4->{\hspace*{0.08cm}{\color{DarkYellow}\underbrace{\begin{pmatrix}y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}_{\bm{b}_{t}}}}$}

\vspace*{-0.6cm}

\only<4>{\hspace*{8cm}\colorbox[RGB]{0,61,102}{$\bY_{t}=\color{orange}\bS\color{DarkYellow}\bm{b}_{t}$}}

\vspace*{10cm}

## Forecasting notation

Let $\hat{\by}_n(h)$ be vector of initial $h$-step forecasts, made at time $n$, stacked in same order as $\by_t$. \pause\newline  (In general, they will not ``add up''.)\pause

\begin{block}{}
Reconciled forecasts must be of the form:
$$\tilde{\by}_{n}(h)=\bS\bm{G}\hat{\by}_{n}(h)$$
for some matrix $\bm{G}$.
\end{block}\pause

 * $\bm{G}$ extracts and combines base forecasts $\hat{\by}_{n}(h)$ to get bottom-level forecasts.
 * $\bS$ adds them up

## Optimal combination forecasts
\fontsize{14}{15}\sf

\begin{alertblock}{Main result}
The best (minimum sum of variances) unbiased forecasts are obtained when
$\bm{G} = (\bS'\bSigma^{-1}_{h}\bS)^{-1}\bS'\bSigma^{-1}_{h}$,
where $\bSigma_h$ is the $h$-step base forecast error covariance matrix.
\end{alertblock}

\pause

\begin{block}{}
\centerline{$\displaystyle\textcolor{red}{\tilde{\by}_{n}(h)}
=\bS(\bS'\bSigma^{-1}_{h}\bS)^{-1}\bS'\bSigma^{-1}_{h}\textcolor{blue}{\hat{\by}_{n}(h)}$}
\end{block}\fontsize{14}{15}\sf

\alert{\textbf{Problem:}} $\bSigma_h$ hard to estimate, especially for $h>1$.

\alert{Solutions:}\vspace*{-0.4cm}

 * Ignore $\bSigma_h$ (OLS)
 * Assume $\bSigma_h$ diagonal (WLS) [Default in `hts`]
 * Try to estimate $\bSigma_h$ (GLS)

## Features
\fontsize{15}{17}\sf

 * Covariates can be included in initial forecasts.
 * Adjustments can be made to initial forecasts at any level.
 * Very simple and flexible method. Can work with *any* hierarchical or grouped time series.
 * Conceptually easy to implement: regression of base forecasts on structure matrix.

## Australian tourism

\full{regions1_with_labels}
\only<3-4>{\begin{textblock}{6.8}(2,2)
\begin{block}{Hierarchy:}
  \begin{itemize}
      \item  States (7)
      \item  Zones (27)
      \item  Regions (82)
  \end{itemize}
\end{block}
\end{textblock}}
\only<4>{\begin{textblock}{6.8}(2,6)
\begin{block}{Base forecasts}
ETS (exponential smoothing) models
\end{block}\end{textblock}}

\only<2>{\begin{textblock}{10.}(1.4,2)
\begin{block}{Domestic visitor nights}
Quarterly data: 1998 -- 2006.\\
From: \textit{National Visitor Survey}, based on annual interviews of 120,000 Australians aged 15+, collected by Tourism Research Australia.
\end{block}
\end{textblock}}

## Base forecasts

\only<1>{\full{austourism1}}
\only<2>{\full{austourism2}}
\only<3>{\full{austourism3}}
\only<4>{\full{austourism4}}
\only<5>{\full{austourism5}}
\only<6>{\full{austourism6}}
\only<7>{\full{austourism7}}
\only<8>{\full{austourism8}}
\only<9>{\full{austourism9}}

## Reconciled forecasts
\only<1>{\full{Australia}}
\only<2>{\full{States}}
\only<3>{\full{Capitals}}

## Forecast evaluation

  * Select models using all observations;
  * Re-estimate models using first 12 observations and generate 1- to 8-step-ahead forecasts;
  * Increase sample size one observation at a time, re-estimate models, generate forecasts until the end of the sample;
  * In total 24 1-step-ahead, 23 2-steps-ahead, up to 17 8-steps-ahead for forecast evaluation.

## Forecast evaluation

```{r rorigin, include=FALSE}
for(i in 2:6)
{
  fname <- paste("rollingorigin",i,sep="")
  savepdf(fname,height=9.5,width=15)
  plot(0,0,xlim=c(0,28),ylim=c(0,1),
       xaxt="n",yaxt="n",bty="n",xlab="",ylab="",type="n")
  for(j in 1:20)
  {
    test <- (6+j):26
    train <- 1:(5+j)
    arrows(0,1-j/20,27,1-j/20,0.05)
    points(train,rep(1-j/20,length(train)),pch=19,col="blue")
    if(length(test) >= i)
      points(test[i], 1-j/20, pch=19, col="red")
    if(length(test) >= i)
      points(test[-i], rep(1-j/20,length(test)-1), pch=19, col="gray")
    else
      points(test, rep(1-j/20,length(test)), pch=19, col="gray")
  }
  text(28,.95,"time")
  endpdf()
}
for(k in 1:20) {
  fname <- paste("rorigin",k,sep="")
  savepdf(fname,height=9.5,width=15)
  plot(0,0,xlim=c(0,28),ylim=c(0,1),
     xaxt="n",yaxt="n",bty="n",xlab="",ylab="",type="n")
  for(j in 1:k) {
    test <- (6+j):26
    train <- 1:(5+j)
    arrows(0,1-j/20,27,1-j/20,0.05)
    points(train,rep(1-j/20,length(train)),pch=19,col="blue")
    if(length(test) >= 1) {
      points(test[1], 1-j/20, pch=19, col="red")
      points(test[-1], rep(1-j/20,length(test)-1), pch=19, col="gray")
    }
    else {
      points(test, rep(1-j/20,length(test)), pch=19, col="gray")
    }
  }
  text(28,.95,"time")
  endpdf()
}
```

\begin{textblock}{14}(0.2,1.4)
\colorbox{white}{\hspace*{0.2cm}\textbf{\textcolor{blue}{Training sets}} \hspace*{3cm}
\textbf{\textcolor{red}{Test sets
\only<1-20>{$h=1$}%
\only<21>{$h=2$}%
\only<22>{$h=3$}%
\only<23>{$h=4$}%
\only<24>{$h=5$}%
\only<25>{$h=6$}%
}}\hspace*{3.22cm}}
\end{textblock}

\only<1>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin1}}
\only<2>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin2}}
\only<3>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin3}}
\only<4>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin4}}
\only<5>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin5}}
\only<6>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin6}}
\only<7>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin7}}
\only<8>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin8}}
\only<9>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin9}}
\only<10>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin10}}
\only<11>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin11}}
\only<12>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin12}}
\only<13>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin13}}
\only<14>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin14}}
\only<15>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin15}}
\only<16>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin16}}
\only<17>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin17}}
\only<18>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin18}}
\only<19>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin19}}
\only<20>{\placefig{0.2}{2.0}{width=12.8cm}{rorigin20}}
\only<21>{\placefig{0.2}{2.0}{width=12.8cm}{rollingorigin2}}
\only<22>{\placefig{0.2}{2.0}{width=12.8cm}{rollingorigin3}}
\only<23>{\placefig{0.2}{2.0}{width=12.8cm}{rollingorigin4}}
\only<24>{\placefig{0.2}{2.0}{width=12.8cm}{rollingorigin5}}
\only<25>{\placefig{0.2}{2.0}{width=12.8cm}{rollingorigin6}}

## Hierarchy: states, zones, regions
\fontsize{10}{10.5}\sf\tabcolsep=0.12cm\vspace*{-0.6cm}
\hspace*{-0.2cm}\begin{tabular}{lrrrrrrr}
            & \multicolumn{6}{c}{\bf Forecast horizon} & \\
\textbf{RMSE} & $h=1$  & $h=2$ & $h=3$ & $h=4$ & $h=5$ & $h=6$ & \textbf{Ave}\\
\midrule
\multicolumn{8}{c}{\bf\alert{Australia}}\\
Base   & 1762.04    & 1770.29    & 1766.02     & 1818.82    & 1705.35    & 1721.17    & \bf 1757.28 \\
Bottom & 1736.92    & 1742.69    & 1722.79     & 1752.74    & 1666.73    & 1687.43    & \bf 1718.22 \\
WLS    & 1705.21    & 1715.87    & \hl 1703.75 & 1729.56    & 1627.79    & \hl1661.24 & \bf 1690.57 \\
GLS    & \hl1704.64 & \hl1715.60 & 1705.31     & \hl1729.04 & \hl1626.36 & 1661.64    & \bf \hl 1690.43 \\
\midrule
\multicolumn{8}{c}{\bf\alert{States}}\\
Base   & 399.77     & 404.16     & 401.92      & 407.26     & 395.38     & 401.17     & \bf 401.61 \\
Bottom & 404.29     & 406.95     & 404.96      & 409.02     & 399.80     & 401.55     & \bf 404.43 \\
WLS    & \hl 398.84 & \hl 402.12 & \hl 400.71  & \hl 405.03 & 394.76     & 398.23     & \bf \hl 399.95 \\
GLS    & \hl 398.84 & 402.16     & 400.86      & \hl 405.03 & \hl 394.59 & \hl 398.22 & \bf \hl 399.95 \\
\midrule
\multicolumn{8}{c}{\bf\alert{Regions}}\\
Base   & 93.15      & 93.38      & 93.45       & 93.79      & 93.50      & 93.56      & \bf 93.47 \\
Bottom & 93.15      & 93.38      & 93.45       & 93.79      & 93.50      & 93.56      & \bf 93.47 \\
WLS    & 93.02      & 93.32      & 93.38       & 93.72      & 93.39      & 93.53      & \bf 93.39 \\
GLS    & \hl 92.98  & \hl 93.27  & \hl 93.34   & \hl 93.66  & \hl 93.34  & \hl 93.46  & \bf \hl 93.34
\end{tabular}
