---
title: "High dimensional time series analysis"
date: "robjhyndman.com/hdtsa"
author: "3. Time series features"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: ../header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 60)
library(fpp3)
library(broom)
```

# Features

## Feature extraction and statistics

\fontsize{9}{10}\sf

```{r features}
tourism %>% features(Trips, feat_stl)
```

## Feature extraction and statistics
\fontsize{8}{9}\sf

```{r features-plot, fig.height=4.6}
tourism %>% features(Trips, feat_stl) %>%
  ggplot(aes(x=trend_strength, y=seasonal_strength_year, col=Purpose)) +
    geom_point() + facet_wrap(vars(State))
```

\only<2->{\begin{textblock}{4.7}(7.8,7.3)
\begin{alertblock}{}\fontsize{10}{10}\sf
\begin{itemize}\tightlist
\item Holidays more seasonal than other travel.
\item WA has strongest trends.
\end{itemize}
\end{alertblock}\end{textblock}}

## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most seasonal time series:

```{r extreme}
most_seasonal <- tourism %>%
  features(Trips, feat_stl) %>%
  filter(seasonal_strength_year == max(seasonal_strength_year))
```

\pause\vspace*{-0.3cm}

```{r extreme2, fig.height=3.}
tourism %>%
  right_join(most_seasonal, by = c("State","Region","Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) + geom_line() +
  facet_grid(vars(State,Region,Purpose))
```

```{r pca, echo=FALSE}
# Save pdf figures
savepdf <- function(file, width=16, height=10) {
  fname <<- paste("figs/",file,".pdf",sep="")
  pdf(fname, width=width/2.54, height=height/2.54, pointsize=10)
  par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
}
endpdf <- function() {
  crop::dev.off.crop(fname)
}
# Compute features
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs="feasts"))
# Compute PCs
pcs <- tourism_features %>% select(-State, -Region, -Purpose) %>%
  prcomp(scale=TRUE) %>% augment(tourism_features)
# Save some PC plots
savepdf("pca1", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
savepdf("pca2", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
savepdf("pca3", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
# Find outliers
outliers <- pcs %>%
  filter(.fittedPC1 > 12 | (.fittedPC1 > 10 & .fittedPC2 > 0))
savepdf("pca4", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1) +
  geom_point(data=outliers, aes(x=.fittedPC1, y=.fittedPC2), col="black", shape=1, size=3)
endpdf()
```

## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r tourismfeatures}
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs="feasts"))
```

```{r echo=FALSE}
tourism_features
```

\begin{textblock}{2.5}(9.6,1.3)
\begin{alertblock}{}\fontsize{10}{12}\sf
All features from the feasts package
\end{alertblock}
\end{textblock}


## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r pcatable}
pcs <- tourism_features %>% select(-State, -Region, -Purpose) %>%
  prcomp(scale=TRUE) %>% augment(tourism_features)
```

```{r echo=FALSE}
pcs
```

\begin{textblock}{2.5}(9.6,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca1}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf


\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca2}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf


\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
```

\only<1>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca3}}
\only<2>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca4}}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r outliers2, fig.height=3.9}
outliers %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
    geom_line() +
    facet_grid(vars(State,Region,Purpose)) +
    ggtitle("Outlying time series in PC space") +
    theme(legend.position = "none")
```

## Acknowledgements


\begin{block}{}
\fontsize{11}{11}\sf
\centering\begin{tabular}{l@{\hspace*{1cm}}l}
\includegraphics[height=4cm, width=10cm]{mitch2} &
\includegraphics[height=4cm, width=10cm]{earowang} \\
Mitchell O'Hara-Wild &
Earo Wang \\
\end{tabular}
\end{block}

\vspace*{1cm}

\begin{alertblock}{}
\centerline{\textbf{feasts.tidyverts.org}}
\centerline{\textbf{robjhyndman.com}}
\end{alertblock}

\vspace*{10cm}